NumPy---Array Programming in Python

### 2020-01-26 body.tex
Minor edits

Users predominantly interact with NumPy arrays using {\em indexing} (to access
subarrays or individual elements), {\em operators} (e.g., $+$, $-$, $\times$,
and, for matrix multiplication, $@$), as well as {\em array-aware functions};
together, these provide an easily readable, expressive, high-level API for
array programming, while NumPy
deals with the underlying mechanics of making operations fast.

@@

The expressive power and brevity of NumPy's array programming syntax combined with 
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa} make interactive data exploration fluid and simple.
It is common for NumPy users to work interactively with their data,
discovering through experimentation bits of code that they stich together into
imperative or functional programs.

@@

While NumPy provides a solid foundation for array programming, the scientific
Python ecosystem is composed of many different libraries, which build on and
extend this functionality \fixme{(Fig.)}.
SciPy provides fundamental algorithms for scientific computing on arrays
including many mathematical, scientific, and engineering routines.
SciPy and NumPy are closely related and were developed together often
by the same people.
Matplotlib generates publication-ready figures and visualizations.
Employing these core array-aware libraries are widely used
technique-specific libraries\cite{SciPyProceedings_11}.
Finally, several domain-specific
libraries \cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,2018EGUGA..2012146H}
are built on this diverse foundation of tooling.

Exposing array programming primitives, as well as the surrounding ecosystem of
tools, in Python---an interpreted language---creates an ideal environment for
interactive, exploratory data analysis where users may iteratively inspect,
manipulate, and visualize their data.
This has made the scientific Python ecosystem popular for scientific
research.

@@

For example, the \code{eht-imaging} library developed by the Event Horizon Telescope 
collaboration relies on many components of the scientific Python ecosystem.
NumPy arrays were used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image 
reconstruction.
Scipy was used to provide tools used in general image processing such as 
filtering and image alignment, while scikit\_image provided higher-level
functionality such as edge filters and Hough transforms.
Other components of Scipy are used throughout the library, such as the
\code{scipy.optimize} package for handling general optimization tasks.
NetworkX \cite{SciPyProceedings_11} was used for image comparisons.
Astropy \cite{astropy:2013, astropy:2018} is used in quite a few places,
primarily for I/O and time/coordinate transformations.
Matplotlib is used for visualizing data throughout the analysis pipeline,
including the generation of the final image of the black hole.
% N.B. - most of this comes from introspecting the Image class defined in
% ehtim/image.py.

@@

NumPy has subsequently seen prolific adoption, with millions of users,
first from research and later also industry.
It spurred the development of the larger scientific Python ecosystem and
heralded the current era of wide-spread use of Python for scientific computing.
The reasons for its success are complex, but include the fact that
NumPy (a) is built on top of Python,
(b) has a strong culture of collaboration and good software development practices, and
(c) provides a cutting-edge array programming environment for scientific data analysis.

>>>>

Users interact with NumPy arrays using {\em indexing} (to access
subarrays or individual elements), {\em operators} (e.g., $+$, $-$, $\times$, and $/$
for elementwise operations and $@$ for matrix multiplication), as well as {\em array-aware functions};
together, these provide an easily readable, expressive, high-level API for
array programming, while NumPy
deals with the underlying mechanics of making operations fast.

@@

@@

The expressive power and brevity of NumPy's array programming syntax combined with 
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa} makes interactive data exploration fluid and simple.
It is common for NumPy users to work interactively with their data,
discovering through experimentation bits of code that they stich together into
imperative or functional programs.
Exposing array programming primitives, as well as the surrounding ecosystem of
tools, in Python---an interpreted language---creates an ideal environment for
interactive, exploratory data analysis where users may iteratively inspect,
manipulate, and visualize their data.

The scientific Python ecosystem \fixme{(Fig.)} is composed of many different libraries.
SciPy provides fundamental algorithms for scientific computing on arrays
including many mathematical, scientific, and engineering routines.
SciPy and NumPy are closely related and were developed together often
by the same people.
Matplotlib generates publication-ready figures and visualizations.
Employing these core array-aware libraries are widely used
technique-specific libraries\cite{SciPyProceedings_11}.
Finally, several domain-specific
libraries \cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,2018EGUGA..2012146H}
are built on this diverse foundation of tooling.
This has made the scientific Python ecosystem popular for scientific
research.

@@

For example, the \code{eht-imaging} library developed by the Event Horizon Telescope 
collaboration relies on many components of the scientific Python ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image 
reconstruction.
Scipy is used to provide tools used in general image processing such as 
filtering and image alignment, while scikit-image provided higher-level
functionality such as edge filters and Hough transforms.
Other components of Scipy are used throughout the library, such as the
\code{scipy.optimize} package for handling general optimization tasks.
NetworkX \cite{SciPyProceedings_11} is used for image comparisons.
Astropy \cite{astropy:2013, astropy:2018} is used in quite a few places,
primarily for I/O and time/coordinate transformations.
Matplotlib is used for visualizing data throughout the analysis pipeline,
including the generation of the final image of the black hole.
% N.B. - most of this comes from introspecting the Image class defined in
% ehtim/image.py.

@@

NumPy has subsequently seen prolific adoption, with millions of users,
first from research and later also industry.
It spurred the development of the larger scientific Python ecosystem and
heralded the current era of wide-spread use of Python for scientific computing.
The reasons for its success are varied and complex, but include the fact that
NumPy (a) is built on top of Python,
(b) has a strong culture of collaboration and good software development practices, and
(c) provides a cutting-edge array programming environment for scientific data analysis.
